MSSM-Penn-Pitt - ACC and DLPFC - DREAM - FEATURECOUNTS - GRCh38 - CQN - Model Selection

### 2020-04-14 cmc-rnaseq-de/MSSM-Penn-Pitt_ACC_DLPFC_FEATURECOUNTS_GRCh38_CQN_Sparse.Rmd
random sampling of genes

```{r gene.param}
## Get GC content from biomart
# backgroundGenes = data.frame(gene_id = rownames(counts)) %>%
#   dplyr::mutate(id = gene_id) %>%
#   tidyr::separate(id, c('ensembl_gene_id','position'), sep = '\\.')
# 
# # Define biomart object
# mart <- useMart(biomart = "ENSEMBL_MART_ENSEMBL", 
#                 host = "uswest.ensembl.org", # Ensembl Release 99 (January 2020)
#                 dataset = "hsapiens_gene_ensembl")
# 
# # Query biomart
# Ensemble2HGNC <- getBM(attributes = c("ensembl_gene_id", "hgnc_symbol", "percentage_gene_gc_content", "gene_biotype", "chromosome_name"),
#                        filters = "ensembl_gene_id", values = backgroundGenes$ensembl_gene_id,
#                        mart = mart)
# 
# GENE.GC.CONT = Ensemble2HGNC %>%
#   dplyr::left_join(backgroundGenes) %>% 
#   dplyr::select(gene_id, percentage_gene_gc_content, chromosome_name) %>%
#   unique
# rownames(GENE.GC.CONT) = GENE.GC.CONT$gene_id]
GENE.GC.CONT = downloadFile("syn21907981")
```

@@

### Library Normalisation
Initial library normalisation usign cqn
```{r cqn}
# Compute offset for gene length and gc content
gene_lengths = tibble::column_to_rownames(gene_lengths, var = 'gene_id') %>% as.data.frame()
CQN.GENE_EXPRESSION = cqn(NEW.COUNTS, 
                          x = GENE.GC.CONT[PROCESSED_COUNTS$filteredExprMatrix$genes$genes, 'percentage_gene_gc_content'],
                          lengths = gene_lengths[PROCESSED_COUNTS$filteredExprMatrix$genes$genes, 'Length'],
                          lengthMethod = "smooth", 
                          verbose = FALSE)
CQN.GENE_EXPRESSION$E = CQN.GENE_EXPRESSION$y + CQN.GENE_EXPRESSION$offset
```

@@

>>>>

```{r gene.param}
# Get GC content from biomart
backgroundGenes = data.frame(gene_id = rownames(counts)) %>%
  dplyr::mutate(id = gene_id) %>%
  tidyr::separate(id, c('ensembl_gene_id','position'), sep = '\\.')

# # Define biomart object
# mart <- useMart(biomart = "ENSEMBL_MART_ENSEMBL", 
#                 host = "uswest.ensembl.org", # Ensembl Release 99 (January 2020)
#                 dataset = "hsapiens_gene_ensembl")
# 
# # Query biomart
# Ensemble2HGNC <- getBM(attributes = c("ensembl_gene_id", "hgnc_symbol", "percentage_gene_gc_content", "gene_biotype", "chromosome_name"),
#                        filters = "ensembl_gene_id", values = backgroundGenes$ensembl_gene_id,
#                        mart = mart)
Ensemble2HGNC = downloadFile("syn21907981")
GENE.GC.CONT = Ensemble2HGNC %>%
  dplyr::left_join(backgroundGenes) %>%
  dplyr::select(gene_id, percentage_gene_gc_content, chromosome_name) %>%
  unique
rownames(GENE.GC.CONT) = GENE.GC.CONT$gene_id
```

@@

### Library Normalisation
Initial library normalisation usign cqn
```{r cqn}
# Compute offset for gene length and gc content
gene_lengths = tibble::column_to_rownames(gene_lengths, var = 'gene_id') %>% as.data.frame()
CQN.GENE_EXPRESSION = cqn::cqn(NEW.COUNTS, 
                          x = GENE.GC.CONT[PROCESSED_COUNTS$filteredExprMatrix$genes$genes, 'percentage_gene_gc_content'],
                          lengths = gene_lengths[PROCESSED_COUNTS$filteredExprMatrix$genes$genes, 'Length'],
                          lengthMethod = "smooth", 
                          verbose = FALSE)
CQN.GENE_EXPRESSION$E = CQN.GENE_EXPRESSION$y + CQN.GENE_EXPRESSION$offset
```

@@

### Subset Genes
```{r subset-genes}
rand_sample <- sample(1:dim(NEW.COUNTS)[1], size = round(dim(NEW.COUNTS)[1]*.6), replace = FALSE)
subset_NEW.COUNTS<- NEW.COUNTS[rand_sample,]

subset_CQN <- CQN.GENE_EXPRESSION$E[rand_sample,]
```
